---
title: "Exploratory Data Analysis"
author: "Bradley Deku"
date: "2023-01-31"
output: word_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


##R Packages
```{r}
library(tidyverse)
library(lessR)

#descr package 
library(descr)
#stats library
library(stats)
```


##Datasets
```{r}

train_labels <- read.csv("C:\\Users\\Brad\\OneDrive - Ashesi University\\Desktop\\Ashesi University\\Senior Year Sem 2\\Capstone 2\\Data\\train_labels.csv")
train_values <- read.csv("C:\\Users\\Brad\\OneDrive - Ashesi University\\Desktop\\Ashesi University\\Senior Year Sem 2\\Capstone 2\\Data\\train_values.csv")

train_data <- merge(train_labels,train_values, by.x = "building_id", by.y = "building_id")

write.csv(train_data, "C:\\Users\\Brad\\OneDrive - Ashesi University\\Desktop\\Ashesi University\\Senior Year Sem 2\\Capstone 2\\Data\\train_data_2.csv", row.names = TRUE)
```



```{r}

head(train_labels)
head(train_values)
head(train_data)
```


```{r}

train_data %>%
  select(building_id,count_floors_pre_eq,age,height_percentage, damage_grade) %>%
  summary()

```
The mean  damage grade in 2.238 /3, which means that majority of buildings suffered a damage grade of 2 or 3.

The mean age of buildings was 26.54 years. The oldest building was 995 years old, adn the newest building was less than a year old.

Buildings had an average of 2.13 or approximately 2 floors. The building with the most floors had 9 floors. 



#Data Management and Subsetting
```{r}

#Grouping number of floors from 1 - 3, where 1 represents 1 floor, 2 =  2 floors and 3 = more than 2 floors.
#This is done since most buildings in the dataset have 2 floors

train_data$floors <- rep(NA, nrow(train_data))
train_data$floors[train_data$count_floors_pre_eq == 1]<- 1
train_data$floors[train_data$count_floors_pre_eq == 2]<- 2
train_data$floors[train_data$count_floors_pre_eq == 3 | train_data$count_floors_pre_eq == 4 | train_data$count_floors_pre_eq == 5 | train_data$count_floors_pre_eq == 6 | train_data$count_floors_pre_eq == 7 | train_data$count_floors_pre_eq == 8 | train_data$count_floors_pre_eq == 9]<- 3
freq(train_data$floors)



#Grouping Buildings according to age (A = less than 15, B = 15-30, C = 31-50, D = older than 50)
train_data$age_range <- rep(NA, nrow(train_data))
train_data$age_range <- as.factor(ifelse(train_data$age<15, 'A', ifelse(train_data$age<31,'B', ifelse(train_data$age<51,'C',ifelse(train_data$age<1000,'D','E' )))))
freq(train_data$age_range)






```

```{r}

Histogram(damage_grade, data = train_data)

hist(train_data$damage_grade)

```

Most buildings were damaged to a grade of 2, followed by 3 and then 1. The training data is skewed towards buildings damaged to a grade of 2.


```{r}

```



























